Computing likelihoods under the SMC for a general class of ARGs.

### 2023-12-05 paper.tex
add figures
supplementary figures

\documentclass{article}
\usepackage[round]{natbib}
\usepackage{amsmath,amssymb,amsfonts, bbm}%
\usepackage{geometry}%
\usepackage{color}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{nameref}
\usepackage[right]{lineno}
\usepackage{subcaption}
\usepackage{tikz}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\textwidth]{figures/smc_custom_2rows_area.png}
\caption{ARG generated under the SMC and its representation as a 
series of local trees (bottom). To ensure the one-to-one correspondance 
between both representations a node is recorded for each local tree whenever 
a node is encoundered along its ancestry path in the graph representation. 
This results in nodes that (locally) have only a single child, and are therefore 
(locally) unary. Blue (unary) nodes represent recombinant nodes. 
The blue arrows shows the familiar left-to-right logic of the SMC whereby 
the floating lineage coalesces randomly
with a the remaining portion of the local tree following recombination.
Note how the presence of the unary node on a local tree is 
indicative of either a past (to the left of the local tree) 
or future (to the right) coalescence event. 
Both ARG representations can be simplified by removing the recombinant nodes. 
In that case the information we have on the time to each of these events is 
limited to the windows (blue shaded array) delimited by by the age of 
the child node $c$ associated with that lineage and the minimum age of 
the parents of the edges connecting to $c$. 
 [TO DO: wanted to use the 
added in haplotype information to explain why we can actually \textit{infer} the 
extent of each lineage, and thus create unary nodes in the first place.]
}
\label{fig:smc-unary}
\end{figure}

@@

All scripts and data used for this manuscript are available on https://github.com/gertjanbisschop/smc-bit-paper
An implementation of the algorithm is available on https://github.com/gertjanbisschop/runsmc.

\bibliographystyle{plainnat}
\bibliography{paper.bib, temp.bib}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.75\textwidth]{figures/supplementary-figs/argweaver_vs_runsmc.png}
\caption{Compare likelihoods inferred ARGs ARGweaver: placeholder for final figure with similar content}
 \label{sup:fig:vs-argweaver}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=0.75\textwidth]{figures/supplementary-figs/v_hudson_unary_simpl.png}
\caption{Compare likelihoods Hudson and SMC with and without unary nodes: placeholder for final figure with similar content}
\label{sup:fig:vs-hudson}
\end{figure}

>>>>

\documentclass{article}
\usepackage[round]{natbib}
\usepackage{amsmath,amssymb,amsfonts, bbm}%
\usepackage{geometry}%
\usepackage{color}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{nameref}
\usepackage[right]{lineno}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{placeins}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\textwidth]{figures/smc_custom_2rows_area.png}
\caption{ARG generated under the SMC and its representation as a 
series of local trees (bottom). To ensure the one-to-one correspondance 
between both representations a node is recorded for each local tree whenever 
a node is encoundered along its ancestry path in the graph representation. 
This results in nodes that (locally) have only a single child, and are therefore 
(locally) unary. Blue (unary) nodes represent recombinant nodes. 
The blue arrows shows the familiar left-to-right logic of the SMC whereby 
the floating lineage coalesces randomly
with a the remaining portion of the local tree following recombination.
Note how the presence of the unary node on a local tree is 
indicative of either a past (to the left of the local tree) 
or future (to the right) coalescence event. 
Both ARG representations can be simplified by removing the recombinant nodes. 
In that case the information we have on the time to each of these events is 
limited to the windows (blue shaded array) delimited by by the age of 
the child node $c$ associated with that lineage and the minimum age of 
the parents of the edges connecting to $c$. 
 [TO DO: wanted to use the 
added in haplotype information to explain why we can actually infer the 
extent of each lineage, and thus create unary nodes in the first place.]
}
\label{fig:smc-unary}
\end{figure}

@@

All scripts and data used for this manuscript are available on https://github.com/gertjanbisschop/smc-bit-paper
An implementation of the algorithm is available on https://github.com/gertjanbisschop/runsmc.
\FloatBarrier
\bibliographystyle{plainnat}
\bibliography{paper.bib, temp.bib}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=0.75\textwidth]{figures/supplementary-figs/argweaver_vs_runsmc.png}
\caption{Simulated 10kb sequence under human-like parameters, 10 diploid samples. Inferred ARG using ARGweaver. Scatterplot is comparison of prior (discretized SMC) as reported by ARGweaver (x-axis) and likelihood as defined here (y-axis).}
 \label{sup:fig:vs-argweaver}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=0.75\textwidth]{figures/supplementary-figs/v_hudson_unary_simpl.png}
\caption{Compare likelihoods Hudson (full ARG) and SMC with and without unary nodes (without recombination nodes): placeholder for final figure with similar content.}
\label{sup:fig:vs-hudson}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=0.75\textwidth]{figures/supplementary-figs/without_unary.png}
\caption{Simulated 1Mb of sequence under human-like parameters. Computed likelihood as defined here with unary nodes (y-axis), and without, blue dots (x-axis). By limiting the number of recombinations (to 1 event) leading up to a single coalescent node we obtain the values in orange.}
 \label{sup:fig:rec-correction}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=0.75\textwidth]{figures/supplementary-figs/hudson_liks_vs_tsinfer_smc.png}
\caption{Simulated full ARG for 1Mb sequence under human-like parameters (100 diploid samples). Computed likelihood (x-axis) under Hudson. Then inferred ARG with \tsinfer following a random sprinkling of mutations and computed likelihood under SMC (y-axis)}
 \label{sup:fig:hudson-smc}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=0.75\textwidth]{figures/supplementary-figs/single_rep_human_like_1gb.png}
\caption{Single replicate from the dataset simulated above with a 100 random sprinklings of mutations. Likelihood under the SMC, mutation rate $\mu=1.25e-8$ (x-axis), and with mutation rate $4*\mu$ (y-axis). Likelihood for the full ARG under Hudson is indicated with the cross.}
 \label{sup:fig:single-tsinfer}
\end{figure}
